### 2014-04-10 ApTranscriptome_TR.Rmd
Working on manuscript in Rmd format. Knit to md. Then compile to pdf using pandoc. Added citations to bib file.

**April 8, 2014**

@@

```{r setup, echo=FALSE, results='hide', message = FALSE}
# Global settings
options(stringsAsFactors=FALSE)

@@

# T-test comparing standard deviation of expression between colonies
t.test(Ar.int.sd$exp_sd, A22.int.sd$exp_sd)
```

@@

# t-test to compare standard deviation of 'Bimodal' transcripts between colonies
t.test(Ar.bim.sd, A22.bim.sd)
```

@@

# t-test
t.test(varshape.out$s.offset, alternative = "two.sided")

@@

# t-test
t.test(varshape.out$s.slope, alternative = "two.sided")

# curvature = curvature1 - curvature2
varshape.out$curvature <- varshape.out$A22.curve - varshape.out$Ar.curve
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.curvature <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.curvature"] <- (varshape.out[i, "curvature"] / mean(c(varshape.out[i, "A22.curve"], varshape.out[i, "Ar.curve"])))
}
# change 'Inf' values to 0
varshape.out[which(varshape.out$Ar.curve == 0 & varshape.out$A22.curve == 0), "s.curvature"] <- 0
# t-test
t.test(varshape.out$s.curvature, alternative = "two.sided")

@@

@@

```{r GO_contingency_test}
responsive.transcripts.ann <- signif.transcripts[names(responsive.transcripts)]
# GO 'response to stress' hits in responsive transcripts
responsive.transcripts.ann[grep("GO:0006950", responsive.transcripts.ann$GO.Biological.Process), list(Transcript, best.hit.to.nr)]

resp.stress.responsive.count <- nrow(responsive.transcripts.ann[grep("GO:0006950", responsive.transcripts.ann$GO.Biological.Process), list(Transcript, best.hit.to.nr)])
# GO 'response to stress' hits in all transcripts
resp.stress.all.count <- nrow(annotationtable[grep("GO:0006950", annotationtable$GO.Biological.Process), list(Sequence.Name, best.hit.to.nr)])

GO.stress.table <- matrix(rbind(resp.stress.responsive.count, nrow(responsive.transcripts.ann) - resp.stress.responsive.count, resp.stress.all.count, nrow(annotationtable) - resp.stress.all.count), nrow = 2)

@@

```{r echo=FALSE, results='hide'}
# combine responsive.transcripts.ann with "type" information
Ap.response.type.sub <- 
responsive.transcripts.ann.type <- merge(responsive.transcripts.ann, Ap.response.type[,c("Transcript", "A22.type", "Ar.type")], by = "Transcript", all = TRUE)
str(responsive.transcripts.ann.type)
write.csv(responsive.transcripts.ann.type, file = "results/Ap_responsive_genes_20140409.csv", row.names = FALSE)
```

@@

```{r grep_shock}
hsp_responsive <- responsive.transcripts.ann[grep("shock", responsive.transcripts.ann$best.hit.to.nr), list(Transcript, best.hit.to.nr)]
nrow(hsp_responsive)

>>>>

**April 10, 2014**

@@

```{r setup, echo=FALSE, results='hide', message = FALSE}
# Global settings
options(stringsAsFactors=FALSE, digits = 3, scipen = 1)

@@

# T-test comparing standard deviation of expression between colonies
t1 <- t.test(Ar.int.sd$exp_sd, A22.int.sd$exp_sd)
```

@@

# t-test to compare standard deviation of 'Bimodal' transcripts between colonies
t2 <- t.test(Ar.bim.sd, A22.bim.sd)
```

@@

# t-test
t3 <- t.test(varshape.out$s.offset, alternative = "two.sided")

@@

# t-test
t4 <- t.test(varshape.out$s.slope, alternative = "two.sided")

# curvature = curvature1 - curvature2
varshape.out$curvature <- varshape.out$A22.curve - varshape.out$Ar.curve
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.curvature <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.curvature"] <- (varshape.out[i, "curvature"] / mean(c(varshape.out[i, "A22.curve"], varshape.out[i, "Ar.curve"])))
}
# change 'Inf' values to 0
varshape.out[which(varshape.out$Ar.curve == 0 & varshape.out$A22.curve == 0), "s.curvature"] <- 0
# t-test
tt.test(varshape.out$s.curvature, alternative = "two.sided")

@@

```{r echo=FALSE, results='hide'}
responsive.transcripts.ann.type <- signif.transcripts[names(responsive.transcripts)]
# combine responsive.transcripts.ann with "type" information
responsive.transcripts.ann.type <- merge(responsive.transcripts.ann, Ap.response.type[,c("Transcript", "A22.type", "Ar.type")], by = "Transcript", all = TRUE)
str(responsive.transcripts.ann.type)
write.csv(responsive.transcripts.ann.type, file = "results/Ap_responsive_genes_20140409.csv", row.names = FALSE)
```

@@

```{r GO_contingency_test}
# GO 'response to stress' hits in responsive transcripts
GO0006950.responsive <- responsive.transcripts.ann.type[grep("GO:0006950", responsive.transcripts.ann.type$GO.Biological.Process), list(Transcript, best.hit.to.nr, A22.type, Ar.type)]

# in high category
GO0006950.responsive[union(with(GO0006950.responsive, grep("High", Ar.type)), with(GO0006950.responsive, grep("High", A22.type))), ]
# in low category
GO0006950.responsive[union(with(GO0006950.responsive, grep("Low", Ar.type)), with(GO0006950.responsive, grep("Low", A22.type))), ]
# in bimodal category
GO0006950.responsive[union(with(GO0006950.responsive, grep("Bimodal", Ar.type)), with(GO0006950.responsive, grep("Bimodal", A22.type))), ]
# in intermediate category
GO0006950.responsive[union(with(GO0006950.responsive, grep("Intermediate", Ar.type)), with(GO0006950.responsive, grep("Intermediate", A22.type))), ]

# Chi-square test to see if 'response to stress' related genes overrepresented in responsive.transcripts compared to full list
resp.stress.responsive.count <- nrow(responsive.transcripts.ann.type[grep("GO:0006950", responsive.transcripts.ann.type$GO.Biological.Process), list(Transcript, best.hit.to.nr)])
# GO 'response to stress' hits in all transcripts
resp.stress.all.count <- nrow(annotationtable[grep("GO:0006950", annotationtable$GO.Biological.Process), list(Sequence.Name, best.hit.to.nr)])

GO.stress.table <- matrix(rbind(resp.stress.responsive.count, nrow(responsive.transcripts.ann.type) - resp.stress.responsive.count, resp.stress.all.count, nrow(annotationtable) - resp.stress.all.count), nrow = 2)

@@

@@

```{r grep_shock}
hsp_responsive <- responsive.transcripts.ann.type[grep("shock", responsive.transcripts.ann.type$best.hit.to.nr), list(Transcript, best.hit.to.nr, A22.type, Ar.type)]
hsp_responsive
